
# Machine Learning Exam 2025

Solution consists of used codes, discussion of results and
visualizations

# ------------------------------------------------------
install.packages("devtools", dependencies = T) 
library(devtools) 
install_version("space", version = "0.1-1.1", repos = "http://cran.us.r-project.org") 
library(space)    # SPACE estimator 

# Libraries
```{r}
if (!require("pacman")) install.packages("pacman")
pacman::p_load(
forecast, # accuracy
glmnet, # Ridge / Lasso
caret, # stratified split, preProcess, train(), confusionMatrix 
devtools, # install_version
space,    # SPACE estimator
igraph,   # network + spectral clustering
Matrix   # sparse matrices etc.
)
```

# ------------------------------------------------------

## Problem 1 - Using college.csv

### 1.A Training a model to predict applications a university receives

```{r Question 1.A}
# Packages 
library(forecast)  #  accuracy() 
library(glmnet)  # Ridge / LASSO 

# ------------------ 0. Load BS pre-process data ------------------

# Reading College-data. We make first column the name for the different rows, as it’s the name of the universities. 
data <- read.csv("College.csv", row.names = 1) 
# The column “Private” is a binary variable, so we turn it into a dummy-variable which R can read 
data$Private <- ifelse(data$Private == "Yes", 1, 0) 

# ------------------ 1. Train/test split (used in both A and B) ------------------

set.seed(1) # setting the seed so we can re-run the code. 
n <- nrow(data) 
train_ind <- sample(1:n, n * 0.6)  # splitting the data into train/test data. Train 60% 
train_data <- data[train_ind, ] 
test_data  <- data[-train_ind, ] 
n_p <- nrow(test_data)   # number of observations in test 

# ------------------ 2.  Predicting APPS ------------------

# Using OLS-model on traindata 
reg_ols <- lm(Apps ~ ., data = train_data) 

summary(reg_ols) 

# Prediction on testdata 
pred_ols <- predict(reg_ols, newdata = test_data) 

# ------------------ 3. Performance measure ------------------

accuracy_ols <- forecast::accuracy(pred_ols, test_data$Apps) 
accuracy_ols 
# Coomparing our test to the naive benchmark 
nb_pred <- rep(mean(train_data$Apps), n_p) 
accuracy_nb <- forecast::accuracy(nb_pred, test_data$Apps) 
accuracy_nb
```

#### Choice of model

The predictor, Apps, is a continuous variable which makes a regression
model useable. With n(777) \> p(17), we have a low-dimensional dataset,
where OLS performs better when the number of predictors is small
relative to the sample size. As the goal is also just prediction, OLS is
a suited model for this problem.

#### Results discussion

The model was trained on 60% of the data, and we see it working
significantly better than the naïve benchmark where we have a smaller
average bias(ME), it’s errors are smaller on average(RMSE) and a much
smaller percentage of bias (MPE). Most critical measure would be
considered to be RMSE, as we want to figure where we find smallest
errors on average. We also compared to the naïve benchmark, just to
showcase it works much better than “your first guess” In the model,
several predictors are highly correlated, which isn’t a problem as we
are only focusing on prediction. When p is small relative to n,
multicollinearity doesn’t really affect predictive performance as much
as it affects interpretability.

## 1.B Training apps again, but using the 17 predictors in addition to their interactions. And comparing with A

```{r Question 1.B}
# ------------------1. Designmatrix with all interactions being (.^2) ------------------ 

#model_matrix removes the intercept column. The function ensures our data is mathematically readable by handling categorical variables and leaving continuous variables as they are. Any non-binary factor variable is converted into a dummy variable. The (.^2) gives us a new column with the interactions of all variables, so A x B, A x C etc. 
X_train <- model.matrix(Apps ~ .^2, data = train_data)[ , -1] 
X_test  <- model.matrix(Apps ~ .^2, data = test_data)[ , -1] 
Y_train <- train_data$Apps 
Y_test  <- test_data$Apps 

# ------------------ 2. Cross-validation for optimal lambda (LASSO: alpha = 1) ------------------ 

cv_out <- cv.glmnet(X_train, Y_train, alpha = 1, nfolds = 5) #cv.glmnet looks through a sequence of possible values for lambda and tells us which one results in the lowest prediction error on new data. We divide it here into 5 randomly equalled folds - considered enough. Eveytime one fold will be set aside as validation set, the remaining 4 will be used to train the model with different values of lambda. The prediction error is calculated by the single validation set. 
plot(cv_out) 
lambda_opt <- cv_out$lambda.min 
lambda_opt 

# ------------------ 3. Final LASSO-model with optimal lambda ------------------ 

reg_lasso <- glmnet(X_train, Y_train, alpha = 1, lambda = lambda_opt) 
coef(reg_lasso)  # looking at all the coefficients, where we see many will be 0, which is a result of the shrinkage. The model forces coefficients to be within a constraint, where our optimal solution is found on this edge where our model touches it. This is the essence of the model as it does feature selection to stay within the constraint. 

# ------------------ 4. Prediction og performance for LASSO ------------------ 

pred_lasso <- as.vector(predict(reg_lasso, newx = X_test)) # converted to vector, as it the standard format in the glmnet package. 
accuracy_lasso <- forecast::accuracy(pred_lasso, Y_test) 
accuracy_lasso
```

#### Choice of model

We chose LASSO because of the shrinkage effect of the model. The
original data itself has 17 predictors, and after the interactions, we
created (17 x 16) / 2 = 136 terms more. With over 150 predictors we
suddenly work in a high dimensional setting, where OLS doesn’t work as
well, causing high variance and overfitting. LASSO adds a penalty (alpha
= 1) that shrinks some coefficients to exactly zero through feature
selection.

#### Results discussion

When we expand the feature space, we see that OLS cannot handle this and
overfits the training data. This is reflected in the test performance,
where RMSE is 1060, where we have 867 for lasso. For regression RMSE is
a good measure because it penalizes larger errors more. But we also see
a better performance in the other metrics. Lasso clearly outperformances
OLS. This is because LASSO can handle the high-dimensional setting
better, by shrinking coefficients and performing feature selection.
LASSO selects only a subset of the predictors, which helps to reduce
overfitting and improve generalization to new data. When looking at the
coefficients of LASSO, we see that many are zero, which is a result of
the shrinkage. The model forces coefficients to be within a constraint,
where our optimal solution is found on this edge where our model touches
it. This is the essence of the model as it does feature selection to
stay within the constraint.


# --------------------------------------------

## Problem 2

### 2.A Which model or models do you think are appropriate for this task? Train a model for the bank and justify your choice of model.

```{r - Question 2.A and 2.B}
# 0. libraries 
library(caret)   # stratified split, preProcess, train(), confusionMatrix 
# 0.1. Load and preprocess data 
bank <- read.csv("UniversalBank.csv") 

# dropping ID-variables (not predictive variables). Zip.code would just make noise in the data unless we chose to distribute in some way, but since it’s not part of the exercise, we simply just remove it. 
bank <- subset(bank, select = -c(ID, ZIP.Code)) 

# We make sure our “goal-variable” is saved as an integer as expects numbers, not text. Afterwards 
# we also create a new column ‘loanfactor’ which takes the numbers 0/1 from personal.loan and 
# makes them categorical with the names no/yes. This is because caret -> confusionmatrix requires 
# input to be factors. So we now have two versions. 
bank$Personal.Loan <- as.integer(bank$Personal.Loan) 
bank$LoanFactor    <- factor(bank$Personal.Loan, levels = c(0, 1), 
labels = c("No", "Yes")) 

# Looking at the Y-distribution 
cat("Outcome distribution (counts):\n") # writing the title and printing numbers of 0 and 1’s 
print(table(bank$Personal.Loan)) 
cat("\nOutcome distribution (proportions):\n") # same as before, just calculated as percentages 
print(prop.table(table(bank$Personal.Loan))) 

# 2. Doing a stratified test/train split 

set.seed(1) # making sure the split is the same everytime we run the code 
train_ind  <- createDataPartition(bank$Personal.Loan, p = 0.6, list = FALSE) # this makes sure that the distribution is the same. So it basically makes sure that if there are 10% “Yes” in the whole data-set, there is also exactly  10% “yes” in the 60% 
train_data <- bank[train_ind, ] 
test_data  <- bank[-train_ind, ] 

# saving the variables in separate vectors 
y_train_num <- train_data$Personal.Loan 
y_test_num  <- test_data$Personal.Loan 
y_train_fac <- train_data$LoanFactor 
y_test_fac  <- test_data$LoanFactor 

# 3. Logistic Regression (baseline model) 

# We use all the predictors, except for the factor-version of the outcome. Predicting personal.loan based on training data 
log_mod <- glm(Personal.Loan ~ . - LoanFactor, 
data   = train_data, 
family = "binomial") 
summary(log_mod) # visualizing 

# Predicted probabilities & classes (cut-off 0.5) on test-data. Type = response makes sure we 
# probabilities as outcome 
prob_log <- predict(log_mod, newdata = test_data, type = "response") 
pred_log_num <- ifelse(prob_log > 0.5, 1, 0) # if probability > 50%, we guess 1(loan) else 0. 
pred_log_fac <- factor(pred_log_num, levels = c(0, 1), labels = c("No", "Yes")) # translating to 0/1 again so that we can compare.  

# Confusion matrix with guess and true + keynumbers. Positive = Yes tells R that the interesting outcome is Yes - meaning they want the loan 
cm_log <- confusionMatrix(pred_log_fac, y_test_fac, positive = "Yes") 
cm_log # printing 
log_err  <- mean(pred_log_num != y_test_num) # calculating errorrate. The mean is the percentage of errors 

log_sens <- cm_log$byClass["Sensitivity"] 
log_spec <- cm_log$byClass["Specificity"]

# 4. k-NN med tuning via k-fold cross-validation (caret::train) 

# Setting predictors - all except outcome and factorversion as we only want input-variables 
X_cols <- setdiff(names(train_data), c("Personal.Loan", "LoanFactor")) 

# making 2 matrices only containing X-variables 
train_X <- train_data[ , X_cols] 
test_X  <- test_data[  , X_cols] 
set.seed(1) # resetting the randomness as CV involves a random division of folds 

# We here use the caret::train with the method KNN 
knn_fit <- train(  
x = train_X,  
# naming the function train, input  
y = y_train_fac,  
method = "knn", 
# output-variable 
preProcess = c("center", "scale"),  
# standardizing all X variables based on 
# train. So we minus all columns with the mean and divide by standard deviation. This makes higher 
# and smaller numbers weight the same, when KNN measures the distance. Family is an example of 
# small number and income high. 
tuneGrid = data.frame(k = 1:15), # meaning we test k = 1, …, 15 
trControl = trainControl(method = "cv", number = 5) 
) # Caret trains the model 5 times for each k (4 train and test on 1), and in the end choses the k with lowest mean CV-error, which also means we don’t need to split into an extra separate validation set 
knn_fit          
# visualization 
k_opt <- knn_fit$bestTune$k 
cat("\nOptimal k chosen by 5-fold CV:", k_opt, "\n\n")  
# Prediction on test-data with optimal/chosen k.  
pred_knn_fac <- predict(knn_fit, newdata = test_X) 
pred_knn_num <- ifelse(pred_knn_fac == "Yes", 1, 0)  
# calculate with the column again. 
# Confusion matrix + keynumbers 
# getting the optimal k and printing it. 
# again converting. So that we can 
cm_knn <- confusionMatrix(pred_knn_fac, y_test_fac, positive = "Yes") 
cm_knn # printing

# saving errorrate, sensitivity and specificity manually  
knn_err  <- mean(pred_knn_num != y_test_num) 
knn_sens <- cm_knn$byClass["Sensitivity"] 
knn_spec <- cm_knn$byClass["Specificity"] 

# 5. Doing a naive rule 

# We first make a list only 0’s as long as y_test_num (tested persons). Which means we here we guess 0=no on all. 
naive_pred_num <- rep(0, length(y_test_num)) 
naive_pred_fac <- factor(naive_pred_num, levels = c(0, 1), 
labels = c("No", "Yes"))  

# chancing no -> 0 
cm_naive <- confusionMatrix(naive_pred_fac, y_test_fac, positive = "Yes") 

# testing how good “always no” strategy is.  
# saving the keynumbers 
naive_err  <- mean(naive_pred_num != y_test_num) 
naive_sens <- cm_naive$byClass["Sensitivity"] 
naive_spec <- cm_naive$byClass["Specificity"] 

# 6. Comparison 

# 6.1 printing the numbers for naïve rule 

cat("Naive rule (always 'No'):\n") 
cat("  Error rate :", round(naive_err, 4), 
" | Sensitivity :", round(naive_sens, 4), 
" | Specificity :", round(naive_spec, 4), "\n\n")

# 6.2 Printing for logistic regression 

cat("Logistic regression:\n") 
cat("  Error rate :", round(log_err, 4), 
" | Sensitivity :", round(log_sens, 4), 
" | Specificity :", round(log_spec, 4), "\n\n") 

# 6.3 Printing for KNN 

cat("k-NN (tuned with 5-fold CV, k_opt =", k_opt, "):\n")
cat("  Error rate :", round(knn_err, 4), 
    " | Sensitivity :", round(knn_sens, 4), 
" | Specificity :", round(knn_spec, 4), "\n") 

```

#### Choice Of Model (2.A)

The goal is to predict if a customer will accept a personal loan, which
is a binary classification problem. So we have to use a supervised
learning algorithm. We have multiple options for such a problem in this
course, which consists of logistic regression, k-NN and the bayes
classifier. We decided to run the logistic model. It models the
probability of acception the loan directly, can handle continuous and
categorical predictors. It also offers interpretable results, which
could provide valuable insights into what drives the classifications of
loan takers.\
We also run the k-NN, which is a non parametric alternative. It uses a
distance based method, so we standardise the predictors and select the
optimal value of k using cross validation.

#### Results Discussion (2.B)

We look at the balance of the Y, and see that it's heavely weighted
towards more 0’s, in this case saying no to the loan. This imbalance
makes the error-rate misleading. Considering the naive rule, it has a
fairly low error rate (0,1025), but with a sensitivity at 0, meaning it
can’t find any customers. So even though it has an error rate at 10,25%,
the model gives no value for the bank. The problem doesnt state anything
about interpretability of what drives predictions, so a solution could
be to run multiple models and choose the best one purely based on
predictive performance. Given the banks goal to find customers who will
accept the loan, we consider sensitivity as the most critical metric. It
measures the ability to identify “yes” customers correctly. The cost of
false negative is considered higher than false positive as false
positive means no possible customer.\
Comparing k-NN and logistic regression, k-NN outperforms. Both are
significantly better than the naive rule, but k-NN achieves a higher
sensitivity and has lowest error-rate. Therefore we also choose k-NN as
a model.

# ----------------------------------------------

## Problem 3 - companies price, dates, info data and Emergency data used

```{r - Loading data and packages for problem 3}
install.packages("devtools", dependencies = T) 
library(devtools) 
install_version("space", version = "0.1-1.1", repos = "http://cran.us.r-project.org") 
library(space)    # SPACE estimator 
library(igraph)   # network + spectral clustering 
library(Matrix)   # sparse matrices etc. 
# (0) Load data and patch missing series with emergency data. Header = false makes sure the first 
# row isnt the name of the columns 
price <- read.csv("companies-price.csv", header = FALSE) 
emerg <- read.csv("emergency-data.csv", header = FALSE)  
dates <- read.csv("companies-dates.csv", header = FALSE)[,1]  # only the first column, which is dates 
info <- read.csv("companies-info.csv", 
header = FALSE, 
stringsAsFactors = FALSE)  # no factors 
# Naming the columns right  
colnames(info) <- c("Ticker", "Name", "Sector") 
# We then replace the first columns in price data as it's all NA - replacing with emergency data 
price[, 1:10] <- as.matrix(emerg) 
# Giving column names from tickers(firms) - easier to interpret 
colnames(price) <- info$Ticker 
```

### 3.A Load in the data and calculate the log returns from the prices for all companies. Multiply the log returns with 100 and use that as your data.

```{r - Question 3.A}
# Converting as matrix and taking the log of all prices 
P <- as.matrix(price) 
logP <- log(P) 
# log returns: r_t = log P_t - log P_{t-1} - which is the formula for log prices 
R_mat <- apply(logP, 2, diff)        
# multiplying the log by 100 
R_mat <- 100 * R_mat 
# Dropping first date - there is no day before the first day, so it's purely NA. Removing it. 
ret_dates <- dates[-1] 
```

### 3.B Use the average log returns as a simple measure of the common factor and filter it out.

```{r - Question 3.B}
# Calculating the meaan for all stocks everyday. This is thé "market" price - removing NA also 
common_factor <- rowMeans(R_mat, na.rm = TRUE) 
# Subtracting the common factor from each stock 
R_demeaned <- R_mat - common_factor 
# This is the data matrix X we will feed into SPACE: 
X <- R_demeaned 
# Cleaning X: dropping data with any NA or zero variance 
na_cols   <- apply(X, 2, function(z) any(is.na(z))) # stocks that have missing data 
zero_var  <- apply(X, 2, function(z) var(z, na.rm = TRUE) < 1e-8) # stocks not moving - variance zero 
drop_cols <- na_cols | zero_var #dropping stocks that are either of the two above 
# removing the columns from drop_cols and save the clean data in X_clean 
cat("Dropping", sum(drop_cols), "columns due to NAs/zero variance\n") 
X_clean   <- X[, !drop_cols] 
t
ickers   <- colnames(X)[!drop_cols] 
n_obs <- nrow(X_clean) 
p_var <- ncol(X_clean) 
cat("Final data matrix: ", n_obs, "obs x", p_var, "stocks\n") # 5030 obs x 83 stocks 
```

### 3.C Use the BIC to select the optimal λ and estimate the network by SPACE. The loop maytake longer to run here than you are used to. Consider candidate values of lambda from 1 to 1.45, with steps of size 0.05.

```{r - Question 3.C}
# Selecting lambda by BIC and estimate network via SPACE 
S_hat <- cov(X_clean)   # calculating sample covariance-matrix 
# reconstructing precision matrix Omega from a space.joint fit 
get_Omega_from_space <- function(fit) { 
Par <- fit$ParCor 
d   <- fit$sig.fit 
p   <- length(d) 
Omega <- matrix(0, nrow = p, ncol = p) 
diag(Omega) <- d 
# Ω_ij = -Par_ij * sqrt(Ω_ii * Ω_jj) 
for (i in 1:p) { 
for (j in 1:p) { 
if (i != j && Par[i, j] != 0) { 
Omega[i, j] <- - Par[i, j] * sqrt(d[i] * d[j]) 
} } } 
return(Omega) } 
# calculating BIC to choose lambda 
bic_space <- function(fit, S, n) { 
  Omega <- get_Omega_from_space(fit) 
  logdet_Om <- as.numeric(determinant(Omega, logarithm = TRUE)$modulus) 
  trace_SO  <- sum(S * Omega) 
  loglik    <- n * (logdet_Om - trace_SO) 
  k <- sum(Omega[upper.tri(Omega)] != 0) 
  bic <- -2 * loglik + k * log(n) 
return(bic) } 

lambda_grid <- seq(1.00, 1.45, by = 0.05) # the values we wanna test 
bic_vals   <- numeric(length(lambda_grid)) 
fits       <- vector("list", length(lambda_grid)) 
parcors    <- vector("list", length(lambda_grid))   # NEW: store ParCor for each lambda 
set.seed(1) 
# making a loop that runs through all lambda values. Lambda is multiplied by n_obs which is required for the space function 
for (i in seq_along(lambda_grid)) { 
  lam <- lambda_grid[i] 
# We use lam1 = lambda * n_obs (T) as in the PS10 example 
  f
  it <- space.joint( 
    X_clean, 
    lam1  = lam * n_obs, 
    lam2  = 0, 
    weight = TRUE) 
  f
  its[[i]]    <- fit 
  parcors[[i]] <- fit$ParCor           
# storing partial correlations 
  bic_vals[i]  <- bic_space(fit, S_hat, n_obs) # storing the BIC score for each model 
  cat("lambda =", lam, ", BIC =", bic_vals[i], "\n") } 
best_idx   <- which.min(bic_vals) # finding the lambda with the lowest BIC-score 
lambda_opt <- lambda_grid[best_idx] 
cat("Optimal lambda by BIC:", lambda_opt, "\n") 
space_fit <- fits[[best_idx]] # getting the winner-model 
Omega_hat <- get_Omega_from_space(space_fit) 
P_hat <- parcors[[best_idx]]

# Making the adjacency matrix. returning true/false - true = 1, false = 0 
A <- 1 * (P_hat != 0) 
diag(A) <- 0 

```

### 3.D Plot the degree distribution of the network.

```{r - question 3.D}
# code 
# Adjacency: edge if Omega_ij != 0, no self-loops 
A <- (Omega_hat != 0) * 1 
diag(A) <- 0 # A is a networkmap with 0 and 1's. But setting the diagonal to 0 
deg <- rowSums(A) #counting connections for each stock - the degree 
# drawing a histogram over degrees 
hist(deg, 
     breaks = 20, 
     main   = "Degree distribution of estimated stock network", 
     xlab   = "Degree", 
     ylab   = "Frequency") 
```

### 3.E Apply spectral clustering to the network. What do you think is a good choice for k here?

```{r - Question 3.E}
g <- graph_from_adjacency_matrix(A, mode = "undirected", diag = FALSE) # converting so that we can use networkfunctions 
deg_vec <- deg 
L <- diag(deg_vec) - A # calculating laplacian matrice 
eig_L <- eigen(L, symmetric = TRUE) # finding smalles eigenvalues 
# Plotting smallest eigenvalues to inspect eigengap 
vals_sorted <- sort(eig_L$values) 
plot(1:10, vals_sorted[1:10], 
     type = "b", 
     xlab = "Index (sorted eigenvalues)", 
     ylab = "Eigenvalue", 
     main = "Smallest eigenvalues of Laplacian\n(eigengap heuristic)") 
# Choosing k =1 (talked about in our answer 
k <- 1    
cat("Using k =", k, "clusters\n") 
# Take eigenvectors for the k smallest non-zero eigenvalues 
ord <- order(eig_L$values) 
U   <- eig_L$vectors[, ord[1:k]] 
set.seed(1) 
cl_kmeans <- kmeans(U, centers = k, nstart = 50)$cluster # using eigenvectors to part stocks in k
groups 
comm <- cl_kmeans
```

#### Answer / plot

The laplacian eigenvalues show that several eigenvalues are extremely
close to 0, indicating many disconnected components and no clear
enginegap. So no meaningful k\>1. We treat the network as having no
nontrivial communities.

### 3.F Plot the network with colours indicating community membership. You do not need to use the layoutByAttribute() function here

```{r - Question 3.F}
set.seed(1) 
lay <- layout_with_fr(g) # Fruchterman-Reingold layout. Placing neighbours close to each other and not neighbours further away. Makes the plot look better 
plot(g, 
layout       
= lay, 
vertex.size  = 4, 
vertex.label = NA, 
vertex.color = comm, 
main = "Estimated stock network\n(colours = spectral clustering communities)")
```

#### Answer / plot

All vertices are the same color because we picked k=1. The plot shows
that most stocks are isolated nodes, and there are a few small connected
components. This matches the structure suggested by the eigen values.

### 3.G Do you see any evidence of the typical network empirical regularities?

```{r - Question 3.G}
cat("Average degree:", mean(deg), "\n") 
cat("Max degree    :", max(deg), " (ticker:", tickers[which.max(deg)], ")\n") 
# Global clustering coefficient (triangles / triplets) 
clust_coeff <- transitivity(g, type = "global") # tells us: if A knows B, B knows C, does A then also know C?  
cat("Global clustering coefficient:", clust_coeff, "\n") 
# Average shortest path length (on largest connected component) 
if (!is.connected(g)) { 
  gcc <- components(g) 
  big_comp <- which.max(gcc$csize) 
  g_big <- induced_subgraph(g, which(gcc$membership == big_comp)) 
  avg_path <- average.path.length(g_big) 
  cat("Average path length (giant component):", avg_path, "\n") 
  } else { 
    avg_path <- average.path.length(g) 
    cat("Average path length:", avg_path, "\n")} 
```

#### Answer

The network is extremely sparse, having removed the common factor. We
don’t see any strong hubs or typical regularities. We spot small
clusters and short paths within the small connected components. It
overall looks fragmented and community structure is very limited.
Overall the empirical patterns appear only in a weak form.


# --------------------------------------------------

## Problem 4

### 4.A How do the three quantities change with the number of predictors?

```{r - Question 4.A and 4.B}
# Setting the pseudo-random number generator seed to 1. 
set.seed(1) 
# Simulation parameters: 
n    <- 50          
n_p  <- 100         
R    <- 1000        
# training size 
# test size 
# number of replications 
p_list <- seq(5, 45, 5) # making a list of numbers from 5-45 with in between. This is the number of predictors we want to test 
# Functions:  
# manually calculating beta-hat.  
basic_ols <- function(Y, X){ 
beta_hat <- solve(t(X) %*% X) %*% (t(X) %*% Y) # the solution to OLS 
return(beta_hat)} 
# simple function where we calculate MSE manually 
mse <- function(Y_hat, Y_true){ 
mean((Y_hat - Y_true)^2) } 
# Storage for results - so that we can easily plot it after. We create three empty matrices with 1000 rows(R) and the amount of columns as values in p_list 
MSE_storage <- matrix(0, nrow = R, ncol = length(p_list)) 
VAR_storage <- matrix(0, nrow = R, ncol = length(p_list)) 
BIAS2_storage <- matrix(0, nrow = R, ncol = length(p_list)) 
colnames(MSE_storage) <- colnames(VAR_storage) <- colnames(BIAS2_storage) <- p_list 
# giving the values from p_list as names for the columns in the other 3 matrices (MSE, VAR, BIAS2). Gives easier interpretation

# Simulation loops 
for (j in seq_along(p_list)) { # first we run a loop through p_list, where P=5 first, after p=10 and so on until 45 
p <- p_list[j] # getting the amount of predictors 
beta_true <- runif(p + 1, min = 0, max = 5) # Draw true beta for this p: length p+1 (incl. intercept), entries ~ U(0,5) which is asked for in the exercise 
# Fixing test design X_test for all replications. Cbind(1) adds a column with 1’s for the intercept.  
X_test <- cbind(1, matrix(rnorm(n_p * p), nrow = n_p, ncol = p))   
f_true_test <- as.vector(X_test %*% beta_true)   # true regression function. This is line we want to hit.  
# Storage of predictions for this p across replications (til bias/var) 
Y_hat_mat <- matrix(0, nrow = n_p, ncol = R) 
# Loop over replications: 
for (r in 1:R) { 
# --- Simulating training data. X_train makes a new training matrix with 50 rows (n). eps_tr creates noise - from rnorm. Y_train creates our Y-values in a vector with the formula Y=Xβ+ϵ 
X_train <- cbind(1, matrix(rnorm(n * p), nrow = n, ncol = p)) 
eps_tr  <- rnorm(n, mean = 0, sd = 1) 
Y_train <- as.vector(X_train %*% beta_true + eps_tr) 
# --- Fit OLS. With the model we try finding β based on training data 
beta_hat <- basic_ols(Y_train, X_train) 
# --- Predicting on test-values. Using the beta_hat values with our X_test to guess Y-values 
Y_hat_test <- as.vector(X_test %*% beta_hat) 
Y_hat_mat[, r] <- Y_hat_test # saving our estimate 
# Simulating noisy test responses to compute MSE: eps_te makes noise for our test-set. Y_test is our true Y in the test set = the true line + the noise, we just created.  
eps_te  <- rnorm(n_p, mean = 0, sd = 1) 
Y_test  <- f_true_test + eps_te 
# Calculating MSE between our guess and the observed test set 
MSE_storage[r, j] <- mse(Y_hat_test, Y_test) 
# calculating the approximate mean of the distances between our guess and the right line.  
VAR_storage[r, j] <- mean((f_true_test - Y_hat_test)^2) 
# Squared bias (OLS almost unbiased here, result therefore very low): 

BIAS2_storage[r, j] <- (mean(f_true_test - Y_hat_test))^2 } } 
# Averaging over replications. We have 1000 results for each p. Here we take the mean of them all, so  we get one value for MSE, VAR and bias^2. 
MSE_mean   <- colMeans(MSE_storage) 
VAR_mean   <- colMeans(VAR_storage) 
BIAS2_mean <- colMeans(BIAS2_storage) 
# simply making a data.frame with the values 
results <- data.frame( 
p      = p_list, 
MSE    = MSE_mean, 
Variance = VAR_mean, 
Bias2    = BIAS2_mean) 
# Plotting the three measures for easier interpretation 
plot(results$p, results$MSE, type = "l", ylim = range(c(MSE_mean, VAR_mean)), 
xlab = "Number of predictors p", ylab = "Value") 
lines(results$p, results$Variance, lty = 2) 
lines(results$p, results$Bias2, lty = 3) 
legend("topleft", 
legend = c("MSE", "Variance", "Bias^2"), 
lty = c(1,2,3), bty = "n")
```

#### Results discussion (4.A)

Variance When looking at the variance, it measures the expected spread
of 𝑓̂(𝑥) around its mean. The quantity here rises with the number of
predictors, meaning it would perform bad on a different set of training
data due to overfitting. As (p) is increasing -\> n the setting comes
closer to a high dimensional setting (when n is fixed), where
overfitting is one of the main problems, causing a very high variance
Squarred bias The squared bias measures the difference between the
expected value of the estimator and the true value. A higher bias would
indicate potential underfitting, but in our model the squared bias
remains close to zero for all values of p. This happens because the data
we generated are from a true linear model and OLS is an unbiased
estimator under the classical assumptions. The expected value of our
true estimator would just be the true function under this assumption,
𝐸[𝑓̂(𝑥)] = 𝑓(𝑥) also meaning 𝑏𝑖𝑎𝑠2 = 0 MSE Looking at the test MSE, Mean
squared error, we see it increasing as p, the number of predictors,
increases. The test MSE is calculated as:

As we concluded the bias is near 0, and the variance is increasing with
p -\> ∞, we only haven’t spoken about the irreducible error (var(ϵ))
which represent the present noise in our data. But this has already been
considered, as we set delta squared = 1 in the simulation, representing
a constant value.\
This gives us a MSE following the variance, but with a higher intercept.

#### Results discussion (4.B)

Some of this has already been considered in (a), but it’s mostly due to
the irreducible error being fixed at 1. This makes MSE 1 higher than the
sum of the squared bias and variance, as MSE follows the variance
otherwise because of our squared bias being near 0. In our simulation we
define the error term as ϵ \~ N(0,1), which is setting the variance of
the noise term to be 1, and as our squared bias is near 0 due to OLS
being an unbiased estimator of our true linear model, the MSE will
approximately be equal to the variance plus the error term, which we
also see being true in the plot.

# ----------------------------------------------------------


## Problem 5

In this problem, we investigate the performance of k-means on a simple
simulated dataset. Create the data in the following way: i) Draw a 200×2
matrix of standard normal values and call it X. ii) Calculate the
Euclidean length of each row of the matrix, and store the result in a
vector. iii) Use the vector of row lengths to find the rows of the
matrix X that have Euclidean length greater than 1, and multiply those
rows with the number 5. Leave the other rows unchanged. iv) Standardise
the matrix X.

```{r - Question 5}
set.seed(1) 
# (i) Drawing a 200 × 2 matrix of standard normal values 
X <- matrix(rnorm(200 * 2), ncol = 2) 
# (ii) Computing Euclidean lengths for each row 
row_len <- sqrt(rowSums(X^2)) 
# (iii) Multiplying rows with length > 1 by 5 
X[row_len > 1, ] <- 5 * X[row_len > 1, ] 
# (iv) Standardise the matrix X (each column) 
X <- scale(X) 
# (iv) Plotting the data  
plot(X, pch = 20, col = "black", 
main = "Simulated data (two clusters)", 
xlab = "X1", ylab = "X2") 
# (iv() Applying k-means 
set.seed(1) 
km <- kmeans(X, centers = 2, nstart = 50) 
# (iv) Plot coloured by cluster 
plot(X, col = km$cluster, pch = 20, 
main = "k-means clustering result", 
xlab = "X1", ylab = "X2")

# Add cluster centres 
points(km$centers, col = 1:2, pch = 8, cex = 2) 
```

#### Results discussion

We see the data form two intuitive clusters, a middle and a cloud
surrounding it, which is further away. This is what happens because
observations with euclidean length \> 1, were multiplied by 5. We apply
the k-means with k=2, and the algorithm does not catch the clusters(ball
and cloud). Instead we see that the outside cloud is divided into 2 and
the middle takes the cluster/color of which of the centroids is closer.
I have plotted the centroid as stars. The result happens because it
assigns each point to the closest centroid, measured by euclidian
distance. It tries to minimize the sum of squared distances within each
cluster, and so we see a different result than what makes intuitive
sense to the naked eye. This illustrates a limitation to the method.
